---
title: "Learning Check 4"
author: "Enter Name Here"
date: "`r format(Sys.time(), '%a/%d/%b')`"
output: 
  html_document:
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages here using library() or p_load() function from {pacman} package

# Import the NERD$ data using the {readxl} package read_xls() function
# Import the BSA data using the {utils} package read.csv() function
```

In our previous Learning Check, we have modeled nested data using an unconditional and conditional fixed effects model. This included a fixed and random intercept. When considering the spaghetti plot you created in Practice2, this was appropriate considering the clear differences in intercepts across school districts. 

**But, what about the differences in slopes between each district?**

## Practice Problem 1

Using the NERD$ data, build a conditional model with grade level ("level") and enrollment ("enroll_raw_NE") as level 1 predictors. Let district_ID be the grouping variable. Create a random intercept and slope for district_ID and only a random slope for enrollment. Do not center the data. 

```{r}
# Conditional model with random slope of enrollment
```

Create the Tau ($\tau$) matrix for the above model. This is also commonly known as the G-matrix in SAS. This will require the `{Matrix}` package. Ensure that the main diagonal of the Tau matrix matches the output from the model's values under its random effects. The off-diagonal represents the covariance between the random effects. 

```{r}
# Tau matrix goes here
```

Using the following equation to calculate the correlation between the random effects:

<p style="text-align: center;"> $\frac{Cov(Intercept, Enrollment)}{sd_{Intercept}+sd_{Enrollment}}$ </p>

```{r}
# Compute random effects correlation by hand
```

## Practice Problem 2

The next model to create involves a cross-level interaction. Because the NERD$ dataset does not naturally include level 2 variables, we will move to using the centered-within-cluster (CWC) age variable that was created in Practice3 from the British Social Attitudes survey. 

Predict support of women's reproductive rights by using the predictors "Sex", "Religion", CWC age, and the interaction of Sex and CWC age. This can be done in two ways: `Sex:CWCAge` or `Sex*CWCAge`. The latter will include Sex, CWCAge, and Sex by CWCAge as separate predictors whereas Sex:CWCAge will only compute one coefficient for the interaction. Include a random intercept and slope for district Also include a random slope for CWCAge. That was a lot of words so here is a list of things to put into the model:

* Sex
* Religion (as a factor)
* CWC Age
* The interaction of Sex and CWC Age
* A random intercept and slope for district_ID
* A random slope for CWC Age

Because we are including both fixed and random effects, we can consider this model a mixed or mixed-effects model!

```{r}
# Create the full model here
```

If Sex is coded as 0 = male and 1 = female, which of the two is associated with greater support for women's reproductive rights?

Which religion category is associated with the greatest support for these rights? For reference:

* Roman Catholic = 1
* Protestant/Church of England = 2
* Other religion = 3
* No religion = 4

Are districts with respondents who are higher than the average age more or less likely to support these rights?

Plot the interaction term involved in the above model using `{ggplot2}`. 

```{r}
# Plot interaction here
```

Visually speaking, do you see existing moderation from the interaction plot?

Using the `lmer` model output, what influence does average age of district have on the association between respondent sex and support of women's reproductive rights? In other words, does the association between sex and support of women's reproductive rights *depend on* the average age of the district that the respondent comes from (or vice-versa)? 

Is this interaction significant?

Decide whether it is worthwhile to retain this interaction by creating the same model but without the interaction term. Then use the `anova()` function (with `refit = F`) to compare model deviances like we did in Practice3.

```{r}
# Create (i.e., copy + paste) the model from above but without interaction
# Compare model deviances
```

## Practice Problem 3

Use the `r2mlm(model)` function to decompose the better fitting model from above. What type of effects make up the majority of the total variance and between-district variance? 

```{r}
# Variance decomposition goes here
```

Using the "Decomposition" estimates, what type of effect (fixed within, fixed between, slope, or mean/intercepts) involves the greatest proportion of variability explained for the total variance?

Using the "Decomposition" estimates, what type of effect (fixed within, fixed between, slope, or mean/intercepts) involves the greatest proportion of variability explained for the within-district variance?

Using the "Decomposition" estimates, what effect (fixed within, fixed between, slope, or mean/intercepts) involves the greatest proportion of variability explained for the between-district variance?