---
title: "Learning Check 2"
author: "Enter Name Here"
date: "`r format(Sys.time(), '%a/%d/%b')`"
output: 
  html_document:
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages here using library() or p_load() function from {pacman} package
# Import NERD$ data ("NE_EdSpending_Jan28th_22") using readxl or similar package
```

## Practice Problem 1

Create a scatterplot of per-pupil expenditure on students ("pp_total_raw_NE") on the y-axis and school enrollment ("enroll_raw_NE") on the x-axis that tracks (i.e., groups) each district's schools by color via the `{ggplot2}` package. This type of plot is sometimes called a "spaghetti plot". Use the below code to filter out school districts with fewer than 2 schools and schools that involve missing data on this variable.

```{r, eval = F}
data_new <- data |> 
  group_by(district_ID) |>
  filter(n() > 1) 

data_new1 <- data_new |> drop_na(pp_total_raw_NE)

data_new1 %>% 
  ggplot(mapping = aes(x = enroll_raw_NE, y = pp_total_raw_NE, group = factor(district_ID))) +
  geom_point() 

# Add geom_smoothing() bit from Ch. 4 of Shaw & Fleck
# Within the geom_smoothing() function, I recommend: `size = .1` and `fullrange = FALSE`

# Insert spaghetti plot here
# Remember to delete eval = F in the chunk header so your code is evaluated
```

Should all school districts be treated as having equal intercepts and slopes? This is how the regression model in Learning Check 1 treated the data -- one intercept and one slope! This does not seem appropriate..

Let's try out our first multilevel model!

## Practice Problem 2 

Using `{lmer}`, ceate a "null", empty, unconditional, or intercept-only model with "pp_total_raw_NE" as the outcome and "district_ID" as the grouping variable. An empty model requires that you have no independent variables predicting the outcome. Here is how this function requires its models to be formatted:

`lmer(DV ~ 1 + IV1 + IV2 + ... + IVp + (random_effect1 + random_effect2 + ... + random_effectp | grouping_variable), data = dataset)`

The initial 1 in the above code is the *fixed* intercept effect. Inputting 1 in place of random_effects serves as the *random* intercept effect. 

```{r}
# Create your empty/null model here
```

How many parameters were estimated (i.e., how many random and fixed effects)? 

What is the fixed effect of the intercept? 

What is the variance of districts around this intercept?

What is the variance of schools around their district's mean?

## Practice Problem 3

Using the output from the null model, answer the following questions:

1. What is the variance attributed to district ($\tau^2_0$)? 

2. What is variance attributed to schools/students ($\sigma^2_0$)?

3. What is the total variance ($\tau^2_0 + \sigma^2_0 = total$)?

4. What is the percent of total variance attributed to district membership?

5. What is the percent of total variance attributed to school membership?

6. Compute the intraclass correlation (ICC) by hand. In laymen terms, this would be the (*variance attributed to district*) / (*total variance*). Here is the equation to do so: 

<p style="text-align: center;"> $\frac{\tau^2_0}{\tau^2_0+\sigma^2_0}$ </p>

Compute the intraclass correlation coefficient (ICC) of the above empty model by using the `{performance}` package.

```{r}
# Compute ICC here
```

Now, use the VarCorr() function from the `{lme4}` package to partition the variance into the Subject (district's) and Residual (school's) variances. This is located at the very bottom of the Whalley text's Variance Partition & ICC module in Ch. 10. 

```{r}
# Partition ICC variances here
```

Do these calculations match between the packages? Do they match your hand calculations?

## Practice Problem 3

Using the null model, insert a level 1 fixed predictor. Insert grade level ("level") as an independent or explanatory variable in the model. If the fixed predictor was a level 2 variable, it would be entered in the same way, but there is not a level 2 predictor of interest in the present data. 

```{r}
# Create fixed factor model
```

A 1-unit increase in grade level is associated with what change in total per-pupil expenditure? Interpret regardless of fixed effect significance. How much variance at level 1 was "soaked up" by including this variable into the model? *Hint:* use the `sigma()` function as found in the example in 5.2.2 of the Shaw & Flake text. 

```{r}
# Compute the variance reduction calculations
```

Calculate the conditional ICC of the new fixed effect model. 

```{r}
# Calculate the ICC of new model here
```

Now that you have entered a fixed effect into the model, reference the conditional (adjusted) ICC. What percent of total per-pupil expenditures is accounted for by grade level?

If you want extra practice, add school enrollment ("enroll_raw_NE") as a level 1 predictor into the conditional model and then re-calculate the reduction in variance and the new conditional ICC. 

```{r}
# Extra practice model
```
